---
title: "After Dada2"
author: "Talia_Karasov"
date: "9/18/2019"
output: html_document
---
The goal of this script is to take the output from dada2 and create filtered ASV tables.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)
library(dada2)
library(dplyr)
library(tidyverse)
#library(fossil)

#library(msa)
#library(DECIPHER)

library(genefilter)
#library(phangorn)
library("RColorBrewer")
#library(gplots)
#library(sjstats)
#library(nlme)

hue1_25 = c("#ada77c","#4477AA", "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#117744", "#44AA77","#114477","#88CCAA", "#777711", "#AAAA44", "#DDDD77", "#774411", "#AA7744", "#DDAA77", "#771122","#DD7788", "darkgoldenrod1", "#771155", "#AA4488", "#CC99BB","#AA7744")
#hue1_25 = c("#bf82cc","#5bc14d","#bf4db3","#9fba36","#7861ce","#4d8c28","#d83e76","#44c181","#d0452f","#4aadd6","#d6812c","#667fc7","#cbaa3b","#9c4769","#7dba6f","#dd809f","#3e8148","#c25d4d","#59c5b0","#de986d","#2f8a72","#91692e","#afb16c","#5f6c2b","#84892d")

```

#######################################################################
# Reading in data and filtering 
#######################################################################
```{r}
#path="/Users/tkarasov/work_main"

#choose whether working on home computer or at work
path="/ebio"
source(paste(path, "/abt6_projects9/pathodopsis_microbiomes/scripts/16S/amp_seq_functions.R", sep=""))

#https://f1000research.com/articles/5-1492/v1 This is the dada2 file

output_direc="/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/"

seqtab.nochim = readRDS(paste(output_direc,"all_runs/demult_python/seqtab_final.rds", sep="/"))

taxa=readRDS(paste(output_direc,"all_runs/demult_python/tax_final.rds", sep="/"))

#metadata=read.table(paste(path,"/abt6_projects9/pathodopsis_microbiomes/pathodopsis_git/data/v1_22_5_merged.txt", sep=""), header=T, sep=",")
#koppen_geiger=read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/pathodopsis_git/data/Pathodopsis_site_metadata_20190808_CZ_course2.txt", header=T, sep="\t")
metadata=read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/pathodopsis_git/data/Pathodopsis_site_metadata_20190808_CZ_course2.txt", header=T, sep="\t")
#metadata=merge(metadata, koppen_geiger, by=c("Site.ID", "Latitude", "Longitude"))
```
Data is read in, now we want to subset it properly to rename the controls (1:26) and give metadata

```{r}
samples.out <- sapply(strsplit(rownames(seqtab.nochim),"_"), "[",1)

#keep the control names
samples.out[1:26]=rownames(seqtab.nochim)[1:26]

rownames(seqtab.nochim) = samples.out

metadata_keep=metadata[metadata$Plant_ID%in%samples.out,]

meta_unique = metadata_keep %>% distinct()

metadata_organized=merge(data.frame(samples.out), meta_unique, by.x="samples.out", by.y="Plant_ID", all.x=TRUE)

subject <- sapply(strsplit(samples.out, "D"), `[`, 1)

samdf <- data.frame(Subject=metadata_organized$samples.out, Latitude=metadata_organized$Latitude, Longitude=metadata_organized$Longitude, Altitude=metadata_organized$Altitude, hpa=metadata_organized$HpA_plant, TourID=metadata_organized$Tour_ID, Clim=metadata_organized$ClimateZ)

rownames(samdf) <- samdf$Subject

sample_names(seqtab.nochim)=samples.out
```

Now let's may the otu_table
```{r}
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(samdf), 
               tax_table(taxa))

rm(seqtab.nochim,taxa)
```

#Pruning
Remove samples with fewwer than 1000 reads
```{r}
GP = prune_samples(sample_sums(ps)>=1000, ps)
GP = prune_samples(sample_data(GP)$TourID!="NA", GP)
rm(ps)
```

Now for taxa names
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(GP))
names(dna) <- taxa_names(GP)
GP <- merge_phyloseq(GP, dna)
taxa_names(GP) <- paste0("seq_", seq(ntaxa(GP)))
rm(dna)
```
FYI: you can access the sequence by refeseq(GP)

Remove mitochondria and samples that have fewer than 50 reads in any sample
```{r}
mito = colnames(otu_table(GP))[which(tax_table(GP)[,5] != "Mitochondria")]
GP = prune_taxa(mito, GP)
flist    <- filterfun(kOverA(1, 50))
GP50 = filter_taxa(GP, flist, TRUE )
GP_fam=tax_glom(GP50, "Family")
#qGPr  = transform_sample_counts(GP, function(otu) otu/sum(otu))
rm(GP)
```

#######################################################################
# Write OTU table
#######################################################################
write fasta file of taxa names
```{r}
taxa_seq = taxa_names(GP)

cat("", file="/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/all_ASVs.fasta", sep='', append=FALSE)
for(i in 1:length(taxa_seq)){
  cat(paste(">seq_",i, sep=""), file="/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/all_ASVs.fasta", sep='\n', append=TRUE)
  cat(taxa_seq[i],file="/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/all_ASVs.fasta", sep='\n', append=TRUE)
}
  
```


#######################################################################
# Basic plot of abundance
#######################################################################


```{r}
#basic plot
top20 <- names(sort(taxa_sums(GP50), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(GP50, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)

plot_bar(ps.top20, x="Subject", fill="Family") + facet_wrap(~Clim, scales="free_x") + theme(axis.line.x = element_blank())

GP50.prop <- transform_sample_counts(GP50, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(GP50.prop, method="NMDS", distance="bray")

plot_ordination(GP50.prop, ord.nmds.bray, color="Clim", title="Bray NMDS") +
  theme_light() +
  scale_fill_manual(values = hue1_25)

nmds = plot_ordination(GP50.prop, ord.nmds.bray, justDF = T)

summary(aov(NMDS1~Clim,data=nmds))

#The richness plotting shouldn't be done with filtered data
plot_richness(GP50.prop, x="Subject", measures=c("Shannon", "Simpson"), color="Clim") +  
  scale_fill_manual(values = hue1_25)
```

# Extract tables

```{r}
extracted_GP50 = as(otu_table(GP50), "matrix")
```

Merge ASV table on genus and family 
```{r}
#GP_genus=tax_glom(GP, "Genus")
```

Rename ASV for family and rename duplicates (but why are there duplicates?)
```{r}
dups = which(duplicated(as.character(tax_table(GP_fam)[,5])))
the_name = paste(as.character(tax_table(GP_fam)[,5])[dups], as.character(dups), sep="_")
tax_table(GP_fam)[,5][dups] = the_name
taxa_names(GP_fam) = as.character(tax_table(GP_fam)[,5])

extracted_GP50_family = as(otu_table(GP_fam), "matrix")
write.table(extracted_GP50_family, "/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/sixteenS_family5.csv", sep=",", col.names = T, row.names=T, quote = F)
```




#make phylogenetic tree. DECIPHER Alignseqs scales linearly rather than exponentially. Cannot be run on my computer. 
#seqs <- getSequences(seqtab.nochim)
#names(seqs) <- seqs # This propagates to the tip labels of the tree
#mult <- AlignSeqs(DNAStringSet(seqs), anchor=NA, verbose=TRUE, processors=16)
#align seqs didn't maintain names

#now write decipher alignment to file 
#https://github.com/benjjneb/dada2/issues/204
#writeXStringSet(mult, file= paste(output_direc,"/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_12_2018/demult_python/16S_otus_aligned.fasta", sep=""))

#fasttree
#system("export OMP_NUM_THREADS=16")
#system("/usr/bin/fasttreeMP -fastest -noml -gtr -nt /ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/16S_otus_aligned.fasta > /ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/16S_otus_aligned.FastTree.tre")

# Alternative for trees...as of right now it seems like we need to use alignseqs to align the OTUs but then put throught raxml
#alignment.rax.gtr <- raxml(DNAbin(mult), m="GTRGAMMAIX", # model f="a", # best tree and bootstrap p=1234, # random number seed x=2345, # random seed for rapid bootstrapping N=100, # number of bootstrap replicates file="alignment", # name of output files exec="raxmlHPC", # name of executablethreads=16S)

#plants not in the metadata  "PA0824" "PA0825" "PA1336" "PA1339" "PA1749" "PA1753" "PA1756" "PA1757" "PA1759" "PA1761" "PC0029"
#st <- seqtab.nochim[rowSums(seqtab.nochim) >= 500,]
#seqtab.nochim <- st
