---
title: "Compare methods"
author: "Talia_Karasov"
date: "9/18/2019"
output: html_document
---
The goal of this script is to take the output from the process_OTU table Rmd and compare the metagenome and 16S data

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(phyloseq)
library("phyloseq")
library(dada2)
library(dplyr)
library(tidyverse)
library(fossil)
library(reshape2)
library(cowplot)
library(intrval)
#library(msa)
#library(DECIPHER)

library(genefilter)
library(phangorn)
library("RColorBrewer")
library(gplots)
library(sjstats)
library(nlme)
library(corrplot)

hue1_25 = c("#ada77c","#4477AA", "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#117744", "#44AA77","#114477","#88CCAA", "#777711", "#AAAA44", "#DDDD77", "#774411", "#AA7744", "#DDAA77", "#771122","#DD7788", "darkgoldenrod1", "#771155", "#AA4488", "#CC99BB","#AA7744")
  
# c("#bf82cc","#5bc14d","#bf4db3","#9fba36","#7861ce","#4d8c28","#d83e76","#44c181","#d0452f","#4aadd6","#d6812c","#667fc7","#cbaa3b","#9c4769","#7dba6f","#dd809f","#3e8148","#c25d4d","#59c5b0","#de986d","#2f8a72","#91692e","#afb16c","#5f6c2b","#84892d")

```

#Comparison between metagenome and 16S classification
Every column is a different plant so we need to normalize by column
```{r}
metagenome_orig2 = read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/metagenome/run116_2018_9_metagenome_reads/meta_family_corrected_per_plant_v2_bacteria.csv", sep=",", header=T, row.names = 1)
#Pseudomonadaceae     0.864007897140246 5.39616182027885e-175

metagenome_orig1 = read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/metagenome/run116_2018_9_metagenome_reads/meta_family_corrected_per_plant.csv", sep=",", header=T, row.names = 1)

#metagenome_orig = read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/metagenome/run116_2018_9_metagenome_reads/centrifuge_output/centrifuge_metagenome_table.txt", sep="\t", header=T, row.names = 1)
#colnames(metagenome_orig)=gsub(".R1.fq.report", "", gsub("..centrifuge_output.", "", colnames(metagenome_orig)))

metagenome1 = sweep(metagenome_orig1, 2, colSums(metagenome_orig1, na.rm=T), FUN="/")
metagenome2 = sweep(metagenome_orig2, 2, colSums(metagenome_orig2, na.rm=T), FUN="/")
reorder_2 = metagenome2[,colnames(metagenome1)]
reorder_1 = metagenome1[rownames(reorder_2),]

###################################################################################################
## COMPARISON BETWEEN OLD AND NEW PIPELINE, 

plot(as.numeric(as.character(unlist(reorder_1))), as.numeric(as.character(unlist(reorder_2))), pch=20)
cor.test(as.numeric(as.character(unlist(reorder_1))), as.numeric(as.character(unlist(reorder_2))))
#metagenome = subset(metagenome, select=-c("family"))
###################################################################################################


sixteenS = as.data.frame(t(read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/sixteenS_family5.csv", sep=",", header=T, row.names = 1)))

sixteenS = sweep(sixteenS, 2, colSums(sixteenS, na.rm=T), FUN="/")

```

Now order them the same
```{r}
keep_id = colnames(metagenome2)[colnames(metagenome2) %in% colnames(sixteenS)]
keep_fam = rownames(metagenome2)[rownames(metagenome2) %in% rownames(sixteenS)]
sixteenS = sixteenS[keep_fam,keep_id]
metagenome2 = metagenome2[keep_fam, keep_id]
#sixteenS = as.matrix(sixteenS)
#metagenome = as.matrix(metagenome)
```

This doesn't look good (rho=0.28). But it seems to be family-specific. For some families we get good correlations but for others not at all
```{r}
cor.test(unlist(metagenome2), unlist(sixteenS))

plot(sqrt(unlist(metagenome2)), sqrt(unlist(sixteenS)), pch = 20, xlab = "Metagenome", ylab = "16S")
corr_mat=as.data.frame(matrix(ncol = 3, nrow=0))
colnames(corr_mat) = c("Family", "correlation", "pval")
for(i in 1:dim(metagenome)[1]){
  if(is.na(metagenome[i,1])!=TRUE){
    curr_cor=cor.test(as.numeric(as.character(metagenome[i,])), as.numeric(as.character(sixteenS[i,])))
    corr_mat[i,]=c(rownames(metagenome)[i], as.numeric(as.character(curr_cor$estimate)), as.numeric(as.character(curr_cor$p.value)))
  }
  else{
    corr_mat[i,]=c(rownames(metagenome)[i], NA, NA)
  }
}

```

Limit to top 10 genera
```{r}
top10_meta = names(sort(rowSums(metagenome_orig), decreasing = T)[1:10])
top10_six = names(sort(rowSums(sixteenS), decreasing = T)[1:10])
u = union(top10_meta, top10_six)
```



```{r}
t_six = data.frame(t(sixteenS))
t_meta = data.frame(t(metagenome2))
t_meta$method = "Metagenome"
t_meta$Plant = rownames(t_meta)
t_six$method = "SixteenS"
t_six$Plant = rownames(t_six)
meta_melt = reshape2::melt(t_meta, variable.name = "Family", value.name = "meta_RA",  idvars = c("Plant", "method"))
six_melt = reshape2::melt(t_six, variable.name = "Family", value.name = "six_RA", idvars = c("Plant", "method"))
all_melt = merge(meta_melt, six_melt, by=c("Plant", "Family"))
all_melt$Family[which(all_melt$Family %ni% u)]=as.factor("Other")


comp = ggplot(data = all_melt, aes(x = (six_RA + 0.0001)*100, y = (meta_RA+0.0001)*100, col = Family )) +
  geom_point() + 
  scale_color_manual(values = hue1_25) +
  theme_bw() +
  geom_abline(slope = 1, linetype = "dashed", color = "Grey", size = 1.0) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("16S RA (%)") +
  ylab("Metagenome RA (%)") +
  theme_minimal_hgrid(12, rel_small = 1) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt"),
    legend.title = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(size=5)))
  
pdf("/ebio/abt6_projects9/pathodopsis_microbiomes/data/figures_misc/metagenome_vs_16S_final.pdf", family = "ArialMT", useDingbats=FALSE)
comp
dev.off()
```

Limit to top 10 genera
```{r}
top10_meta = names(sort(rowSums(metagenome_orig), decreasing = T)[1:10])
top10_six = names(sort(rowSums(sixteenS), decreasing = T)[1:10])
u = union(top10_meta, top10_six)
```

```{r}
metagenome$family = rownames(metagenome)
sixteenS$family = rownames(metagenome)


meta_melt = melt(metagenome, id="family")
meta_melt = meta_melt[meta_melt$family%in%u,]

sixteen_melt = melt(sixteenS, id="family")
sixteen_melt = sixteen_melt[sixteen_melt$family%in%u,]
p_meta = ggplot(data=meta_melt, aes(x = variable, y = value, fill = family)) +
  geom_bar(aes(), stat="identity", position="stack") +
  scale_fill_manual(values = hue1_25)

p_16S = ggplot(data=sixteen_melt, aes(x = variable, y = value, fill = family)) +
  geom_bar(aes(), stat="identity", position="stack") +
  scale_fill_manual(values = hue1_25)

legend1 = get_legend(p_16S)

plot_grid(p_meta + theme(legend.position = "none") , p_16S + theme(legend.position = "none", axis.title.y = element_blank()), legend1, nrow = 1)

pdf("/ebio/abt6_projects9/pathodopsis_microbiomes/data/figures_misc/metagenome_vs_16S_disaster.pdf")
plot_grid(p_meta + theme(legend.position = "none") , p_16S + theme(legend.position = "none", axis.title.y = element_blank()), legend1, nrow = 1)
dev.off()

```


Let's look at raw reads
```{r}
meta_uncorrec = read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/metagenome/run116_2018_9_metagenome_reads/centrifuge_output/centrifuge_metagenome_table_bac.txt", sep="\t", header = T, row.names = 1)
colnames(meta_uncorrec) = gsub(".R1.fq.out", "", colnames(meta_uncorrec))
rownames(meta_uncorrec) = meta_uncorrec$family
meta_uncorrec = subset(meta_uncorrec, select=-c(family))

colSums(meta_uncorrec)
unc_ra = sweep(meta_uncorrec, 2, colSums(meta_uncorrec, na.rm=T), FUN="/")

unc_ra = unc_ra[,colnames(sixteenS)]

shared_fam = intersect(rownames(unc_ra), rownames(sixteenS))

unc_ra = unc_ra[shared_fam,]
sixten_unc = sixteenS[shared_fam,]
plot(unlist(unc_ra), unlist(sixten_unc))
#now combine wwith 16S
```




#comparing filtering
How much does the filtering influence the relative abundances?
```{r}
sixteenS = as.data.frame(t(read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/sixteenS_family.csv", sep=",", header=T, row.names = 1)))
sixteenS_5 = as.data.frame(t(read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/sixteenS_family5.csv", sep=",", header=T, row.names = 1)))

sixteenS = sweep(sixteenS, 2, colSums(sixteenS, na.rm=T), FUN="/")
sixteenS_5 = sweep(sixteenS_5, 2, colSums(sixteenS_5, na.rm=T), FUN="/")
sixteenS_5 = sixteenS_5[rownames(sixteenS),]

#The filtering does not have a strong influence on the relative abundances
plot(as.matrix(sixteenS), as.matrix(sixteenS_5), pch=20, xlab="Relative Abundance, filtering at 50", ylab="Relative Abundance, filtering at 5")
cor.test(as.matrix(sixteenS), as.matrix(sixteenS_5))

```



#Rescaling the microbiome
Let's move on for now and rescale our 16S file with the estimated load from the metagenome calculation
```{r}
met_orig_sub = metagenome_orig[rownames(sixteenS),colnames(subset(sixteenS, select=-c(family)))]
to_transform = subset(sixteenS, select=-c(family))
rescale = sweep(to_transform, 2, colSums(met_orig_sub, na.rm=T), FUN="*")

#Reorder every sample according to load
rescale = rescale[,order(colSums(met_orig_sub, na.rm=T), decreasing = T)]
rescale = rescale[order(rowSums(met_orig_sub, na.rm=T),decreasing = T),]

top10 = names(sort(rowSums(to_transform), decreasing = T)[1:10])
rescale$family=rownames(rescale)
rescale$family[rescale$family%ni%top10]="Other"

#reorder=sort(colSums(met_orig_sub, na.rm=T), decreasing = T)
rescale_melt=melt(rescale, id="family")

p_16S = ggplot(data=rescale_melt, aes(x = variable, y = value, fill = family)) +
  geom_bar(aes(), stat="identity", position="stack") +
  scale_fill_manual(values = hue1_25) +
  xlab("Plants") +
  ylab("Microbial Genome/Plant Genome") +
  theme(axis.text.x=element_blank())

p_16S
```

Who is the most correlated with load?
```{r}
plot(as.numeric(as.character(rescale["Pseudomonadaceae",][1:100])), colSums(met_orig_sub, na.rm=T)[1:100])

corrplot(cor(t(rescale)))
```


What about plotting just the sixteenS data
```{r}
sixteenS = as.data.frame(t(read.table("/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_all/all_runs/demult_python/sixteenS_family5.csv", sep=",", header=T, row.names = 1)))

sixteenS = sweep(sixteenS, 2, colSums(sixteenS, na.rm=T), FUN="/")
top5 = rownames(sixteenS)[order(rowSums(sixteenS), decreasing = T)][1:8]
sixteenS$family = rownames(sixteenS)
sixteenS$family[sixteenS$family %ni% top10]="Other"
#sixt_rescale = sixteenS[,order(sixteenS["Burkholderiaceae",], decreasing = T)]

#reorder the families in the barplot
sixt_rescale$family = fct_relevel(sixt_rescale$family, "Other", after = 0)
sixt_rescale$family = fct_relevel(sixt_rescale$family, "Burkholderiaceae", after = 10)

six_melt = melt(sixt_rescale, id="family")
#six_melt$variable=reorder(six_melt$variable, six_melt$value, function(x) max(table(x)[1]))


p_16S_alone = ggplot(data=six_melt, aes(x = variable, y = value, fill = family)) +
  geom_bar(aes(), stat="identity", position="stack") +
  scale_fill_manual(values = hue1_25) +
  xlab("Plants") +
  ylab("16S relative Abundance") +
  theme(axis.text.x=element_blank())

p_16S_alone
```



















































###################################################
STUFF THAT CAN BE REMOVED LATER
###################################################


#looking at the demultiplexing
So these results are puzzling. First of all, why does the metagenome say there is so much more Burkholderia than does the 16S analysis. Secondly, there is such a poor correlation between these two datasets. And lastly, why do the controls have so many reads.

So why is burkholderia so common? Let's choose a few samples that have a lot of Burkholderia
metagenome_orig["Burkholderiaceae",][which(metagenome_orig["Burkholderiaceae",]>2)]
PA0343   PA0344   PA0367   PA0370   PA0373   PA0376   PA0391   PA0394   PA0397   PA0400   PA0403   PA0406   PA0692   PA0832   PA0842

 PA0895   PA0913   PA1117   PA1119   PA1141   PA1156   PA1179   PA1184   PA1194   PA1196   PA1208   PA1210  PA1214   PA1256   PA1266   PA1268   PA1273


PA0913 has 22947 reads for burkholdereacea in centrifuge_metagenome_table.txt
ST-J00101:116:H23WVBBXY:3:1121:15554:31118 appears five times. It has four of the same taxonomic ranking and one different ranking. 



The following read was assigned to PA0480: 
@HWI-M00333R1:141:000000000-BWDK4:1:2115:11050:21251 1:N:0:CATGGACGTCTTTCCC
GACCTACGGGAGGCAGCAGTGGGGAATTTTCCGCAATGGGCGCAAGCCTGACGGAGCAATGCCGCGTGGAGGTAGAAGGCCTACGGGTCCTGNACTTCTTTTCCCAGAGAAGAAGCAATGACGGTATCTGGGGAATAAGCATCGTCTAACTCTGTGCCAGCAGCCGCGGTAATACAGAGGATGCAAGCGTTATCCGGAATGATTGGGCGTAACGCGTCTGTAGGTGGCTGTTTAAGTCCGCCGTCAACTCCCAGGGCTAAACCCTGGACAGGCGGTGGAAACTACCAAGCTTGAGTACGG

Reverse read:
@HWI-M00333R1:141:000000000-BWDK4:1:2115:11050:21251 2:N:0:CATGGACGTCTTTCCC
TGACTACGGACTACTGGGGTATCTAATCCCATTCGCTCCCCTAGCTTTCGTCTCTCAGTGTCAGTGTCGGCCCAGCAGAGTGCTTTCGCCGTTGGTGTTCTTTCCGATCTCTACGCATTTCACCGCTCCACCGGAAATTCCCTCTGCCCCTACCGTACTCAAGCTTGGTAGTTTCCACCGCCTGTCCAGGGTTGAGCCCTGGGATTTGACGGCGGACTTAAAAACCCCCCTACAGACGTTTTACCCCCAATCCTTCCGGATAACCTTTACCTCACCTTTTTTACCTCGGCTGCTGCCACC

What do I know about this sample?
"CATGGACG TCTTTCCC" corresponds to ???

What went wrong in the demultiplexing here? This reverse reads starts with "TGACTACGGACTAC" so why was it assigned to an index that should be 806.F5?

PA0480		plate7	C7	341F1.806F5
The barcode for position C7 is C7,,,,,CATGGACGC,,AGATCTCG,pathodopsis_16S,TaliaKarasov
F1 = GACCTACGGGAGGCAGCAG
F5(R) = GTAGTCCGTAGTC

if I understand the problem correctly, then these reads should also be in there:
@HWI-M00333R1:141:000000000-BWDK4:1:2119:15080:24654 2:N:0:CATGGACGTCTTTCCC
@HWI-M00333R1:141:000000000-BWDK4:1:2119:15004:24265 2:N:0:CATGGACGTCTTTCCC
@HWI-M00333R1:141:000000000-BWDK4:1:2119:10067:23999 2:N:0:CATGGACGTCTTTCCC

In my demuliplexing from a few months ago this read was place in the PA0480 file:
@HWI-M00333R1:141:000000000-BWDK4:1:1101:10539:5772 1:N:0:CATGGACGTCTTTCCC
reverse read:
GACTACGGACTACTAGGGTTTCTAATCTGGAGTTCAGACGTGCGCTCTTCTGCTGCCTCCCGTAGGTCAGATCGGAAGAGCGTCGTGCAGGGAAAGAGCGTAGAGCTCGGTGGTCGTTGCACCACTAAAAAAAGAATTCCAGTCACCACTCACGAACTCTAGTCCGCCCGTATCAAAGTCAGACCCGTAATTTAGCCCAGTGCTTTTGACTCCGCCGCAACGCACCGACTACTGAGTCATTACGCCCAATACATCATGTCCACGCGACCCCCCTACGTGATACCTACTCTTCTTGCCAGC

This was the proper demultiplexing...I wonder what is going on now...
This directory seems to be oK
/ebio/abt6_projects9/pathodopsis_microbiomes/data/processed_reads/16S/16S_12_2018/demult_python


OK I stripped again. This read was present in the stripped file. and in the all. file
@HWI-M00333R1:141:000000000-BWDK4:1:1101:14324:2477 2:N:0:CATGGACGTCTTTCCC
CCTGTTCGCTCCCCATGCTTTCGCTCCTCAGCGTCAGGTAATGCCCAGAGAACCGCCTTCGCCACCGGTGTTCCTCCTGATATCTGCGCATTTCACCGCTACACCCGGAATTCCGTTCTCCCCTTCATACCTCTAGTCTGCCCGTATCGAAAGCCAGCCAGGTGTTAAGCACCGGGTTTTCACTCCCGACGCGACCAACCGCCTCCGAGCCCTTTACGCCCAATAATTCCGTACCACGCTTGCCCCCTACGTATTACCTCAGCTTCTGGCCCGC



So the same run was demultiplexed incorrectly now...why would this be?
I think the issue was in my clip_barcodes_after_python.sh script. The samples look good until the stripping.